Comb associative operads

### 2018-07-11 CAs_operads.tex
Ajout mini-intro section comb associative operads

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generalizations of the associative operad} \label{sec:CAs_d}
\Todo{Write mini intro.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comb associative operads}
\Todo{Write mini intro.}

It is known that the rewrite rule $\Rew$
orienting~\eqref{equ:congruence_as} from left to right is a convergent
orientation of (\ref{equ:congruence_as}). Then, a Poincaré-Birkhoff-Witt
basis of $\As$ is the set of all right comb binary trees.
\Todo{SG: to maybe to move at another place. }
\medbreak

Let, for any $\gamma \geq 1$, the binary trees
\begin{equation}
    \LComb{\gamma} :=
    \underbrace{(\dots (\Product \circ_1 \Product) \circ_1 \dots)
        \circ_1 \Product}
    _{\gamma \mbox{ \footnotesize operands}}
\end{equation}
and
\begin{equation}
    \RComb{\gamma} :=
    \underbrace{\Product \circ_2
        (\dots \circ_2 (\Product \circ_2 \Product) \dots)}
    _{\gamma \mbox{ \footnotesize operands}}
\end{equation}
be respectively the left and the right comb of degree $\gamma$.
Let us now define for any $\gamma \geq 1$ the \Def{$\gamma$-comb
associative operad} $\CAs{\gamma}$ as the quotient operad
$\Mag/_{\CongrCAs{\gamma}}$ where $\CongrCAs{\gamma}$ is the smallest
operad congruence of $\Mag$ satisfying
\begin{equation} \label{equ:congruence_CAs_gamma}
    \LComb{\gamma} \enspace \CongrCAs{\gamma} \enspace \RComb{\gamma}.
\end{equation}
Notice that $\CongrCAs{1}$ is trivial so that
$\CAs{1} = \Mag$ and that $\CongrCAs{2}$ is the operad congruence
defined by
\begin{equation}
    \begin{tikzpicture}[xscale=.24,yscale=.24,Centering]
        \node(0)at(0.00,-3.33){};
        \node(2)at(2.00,-3.33){};
        \node(4)at(4.00,-1.67){};
        \node[NodeST](1)at(1.00,-1.67)
            {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00)
            {\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(3);
        \node(r)at(3.00,1.5){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \Congr
    \begin{tikzpicture}[xscale=.24,yscale=.24,Centering]
        \node(0)at(0.00,-1.67){};
        \node(2)at(2.00,-3.33){};
        \node(4)at(4.00,-3.33){};
        \node[NodeST](1)at(1.00,0.00)
                {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.67)
                {\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,
\end{equation}
so that $\CAs{2} = \As$.
Let also
\begin{equation}
    \CAsAll := \left\{\CAs{\gamma} : \gamma\geq 1\right\}
\end{equation}
be the set of all the $\gamma$-comb associative operads.
Our purpose is to show that $\CAsAll$ admits a lattice structure. For
that, we provide a description of operads morphisms between the
elements of~$\CAsAll$.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A lattice of operads}
\Todo{Write mini intro.}

@@

\begin{Theorem}\label{thm:lattice_CAs}
    The poset $\left(\CAsAll, \OrdCAs\right)$ is a lattice where
    the lower-bound and upper-bound are, for any positive integers
    $\gamma$ and $\gamma'$, the following operads
    \begin{equation}
      \CAs{\gcd \left(\gamma-1, \gamma'-1\right)}\ \text{and}\
      \CAs{\lcm \left(\gamma-1, \gamma'-1\right)}.
    \end{equation}
\end{Theorem}
\medbreak

@@

We obtain a
new lattice $\left(\KCAsAll,\OrdCAs,\gcd,\lcm\right)$, where $\KCAsAll$
is the set of all operads $\KCAs{\gamma}$. In this linear framework, the
condition $\KCAs{\gamma}\OrdCAs\KCAs{\gamma'}$ means that the dimension
of the space $\text{hom}\left(\KCAs{\gamma'},\KCAs{\gamma}\right)$ is
equal to $1$. Hence, $\left(\KCAsAll,\OrdCAs,\gcd,\lcm\right)$ is related
to the lattice $\left(\QMag,\OrdKMag,\wedge,\vee\right)$ presented in
Section~\ref{sec:Magmatic_operads} by the following proposition.

\begin{Proposition}
  The inclusion
  $\iota:\left(\KCAsAll,\OrdCAs\right)\to\left(\QMag,\OrdKMag\right)$ is
  non decreasing. In particular, for any positive integers $\gamma$ and
  $\gamma'$, we have
  \begin{equation} \label{equ:comparison_of_lattice_operations}
    \KCAs{\gcd \left(\gamma-1, \gamma'-1\right)}\OrdKMag
    \KCAs{\gamma}\wedge\KCAs{\gamma'}
  \end{equation}
  and
  \begin{equation}
    \KCAs{\gamma}\vee\KCAs{\gamma'}\OrdKMag
    \KCAs{\lcm \left(\gamma-1, \gamma'-1\right)}
  \end{equation}

  \end{Proposition}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generalizations of the associative operad}
\label{sec:CAs_d}

In this section, we define comb associative operads and we show that the
set of such operads admits a lattice structure, isomorphic to the lattice
division for nonnegative integers. We also relate this lattice to the
lattice of linear magmatic quotients.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comb associative operads}

Let, for any integer $\gamma \geq 1$, the binary trees
\begin{equation}
    \LComb{\gamma} :=
    \underbrace{(\dots (\Product \circ_1 \Product) \circ_1 \dots)
        \circ_1 \Product}
    _{\gamma \mbox{ \footnotesize operands}}
\end{equation}
and
\begin{equation}
    \RComb{\gamma} :=
    \underbrace{\Product \circ_2
        (\dots \circ_2 (\Product \circ_2 \Product) \dots)}
    _{\gamma \mbox{ \footnotesize operands}}
\end{equation}
be respectively the left and the right comb of degree $\gamma$.
Let us now define for any $\gamma \geq 1$ the \Def{$\gamma$-comb
associative operad} $\CAs{\gamma}$ as the quotient operad
$\Mag/_{\CongrCAs{\gamma}}$ where $\CongrCAs{\gamma}$ is the smallest
operad congruence of $\Mag$ satisfying
\begin{equation} \label{equ:congruence_CAs_gamma}
    \LComb{\gamma} \enspace \CongrCAs{\gamma} \enspace \RComb{\gamma}.
\end{equation}
Notice that $\CongrCAs{1}$ is trivial so that
$\CAs{1} = \Mag$ and that $\CongrCAs{2}$ is the operad congruence
defined by
\begin{equation}
    \begin{tikzpicture}[xscale=.24,yscale=.24,Centering]
        \node(0)at(0.00,-3.33){};
        \node(2)at(2.00,-3.33){};
        \node(4)at(4.00,-1.67){};
        \node[NodeST](1)at(1.00,-1.67)
            {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00)
            {\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(3);
        \node(r)at(3.00,1.5){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \Congr
    \begin{tikzpicture}[xscale=.24,yscale=.24,Centering]
        \node(0)at(0.00,-1.67){};
        \node(2)at(2.00,-3.33){};
        \node(4)at(4.00,-3.33){};
        \node[NodeST](1)at(1.00,0.00)
                {\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.67)
                {\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(3);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,
\end{equation}
so that $\CAs{2} = \As$.
Let also
\begin{equation}
    \CAsAll := \left\{\CAs{\gamma} : \gamma\geq 1\right\}
\end{equation}
be the set of all the $\gamma$-comb associative operads.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Lattice of comb associative operads}

In order to introduce the lattice structure of $\CAsAll$, we study operad
morphisms betweeen its elements by mean of intermediate lemmas.

@@

\begin{Theorem}\label{thm:lattice_CAs}
  The tuple $\LatCAs$ is a lattice, where $\InfCAs$ and $\SupCAs$ are
  defined, for any positive integers $\gamma$ and $\gamma'$, as follows:
    \begin{equation}
      \CAs{\gamma}\InfCAs\CAs{\gamma'}=\CAs{\gcd \left(\gamma-1, \gamma'-1\right)}
    \end{equation}
    and
    \begin{equation}
      \CAs{\gamma}\SupCAs\CAs{\gamma'}=\CAs{\lcm \left(\gamma-1, \gamma'-1\right)}.
    \end{equation}
\end{Theorem}
\medbreak

\begin{Remark}
  The lattice $\left(\N,\mid, \gcd, \lcm\right)$ admits $1$ as minimum
  element and $0$ as maximum element: any nonnegative integer is
  divisible by $1$ and divides $0$. The corresponding property for
  $\LatCAs$ is that it admits $\As=\CAs{2}$ as minimum element and $\Mag=
  \CAs{1}$ as maximum element: any comb associative operad projects onto
  $\As$ and is a quotient of $\Mag$.
\end{Remark}

We terminate this section by relating the lattice
$\LatQMag$ introduced in Section~\ref{sec:Magmatic_operads} to the
lattice $\LatCAs$.

@@

We obtain a new lattice $\LatKCAs$, where $\KCAsAll$ is the set of all
operads $\KCAs{\gamma}$. In this linear framework, the condition
$\KCAs{\gamma}\OrdCAs\KCAs{\gamma'}$ means that the dimension of the
space $\text{hom}\left(\KCAs{\gamma'},\KCAs{\gamma}\right)$ is equal to
$1$. Hence, $\LatKCAs$ is related to $\LatQMag$ by the following
theorem.

\begin{Theorem}
  The inclusion
  $\iota:\left(\KCAsAll,\OrdCAs\right)\to\left(\QMag,\OrdQMag\right)$ is
  non decreasing. In particular, for any positive integers $\gamma$ and
  $\gamma'$, we have
  \begin{equation} \label{equ:comparison_of_lattice_operations}
    \KCAs{\gcd \left(\gamma-1, \gamma'-1\right)}\OrdQMag
    \KCAs{\gamma}\InfQMag\KCAs{\gamma'}
  \end{equation}
  and
  \begin{equation}
    \KCAs{\gamma}\SupQMag\KCAs{\gamma'}\OrdQMag
    \KCAs{\lcm \left(\gamma-1, \gamma'-1\right)}.
  \end{equation}

  \end{Theorem}

\begin{Remark}
  Note that $\LatKCAs$ does not embed into a sub-lattice of $\LatQMag$,
  that is $\iota$ is not a lattice morphism. Consider for instance
  $\gamma=3$ and $\gamma'=4$, then $\KCAs{\gamma}\InfCAs\KCAs{\gamma'}$
  is equal to $\KCAs{2}=\K\As$, whereas
  $\KCAs{\gamma}\InfQMag\KCAs{\gamma'}$ is quotient of $\KMag$ be the two
  relations $\LComb{3}-\RComb{3}$ and $\LComb{4}-\RComb{4}$.
\end{Remark}

### 2018-07-11 Magmatic_operads.tex
Ajout mini-intro section comb associative operads

In this section, we fix a commutative field $\K$ and we study quotients
of the free linear operad over one binary generator. We call this free
operad the \Def{linear magmatic operad} and we denote it by $\KMag$. For
each arity $n$, $\KMag(n)$ is the vector space with basis $\Mag(n)$, and
the compositions of $\KMag$ extend by linearity the ones of $\Mag$.

@@

We denote by $\IMag$ the set of operadic ideals of $\KMag$, and
\begin{equation} \label{equ:definition_of_QMag}
  \QMag =\Big\{\KMag/ I\mid I\in\IMag\Big\},
  \end{equation}
the set of quotients of the linear magmatic operad. Given two elements
$\Oca_1$ and $\Oca_2$ of $\QMag$, we denote by
$\text{hom}\left(\Oca_1,\Oca_2\right)$ the set of operads morphisms from
$\Oca_1$ to $\Oca_2$.

@@

\begin{proof}
  The operad $\Oca_1$ is spanned by the binary generator
  $[\Product]_{I_1}$, which is the class of the element
  $\Product\in\KMag$ in $\Oca_1$. Hence, a morphism
  $\varphi:\Oca_1\to\Oca_2$ is uniquely determined by
  $\varphi\left([\Product]_{I_1}\right)$. Moreover,
  $\varphi\left([\Product]_{I_1}\right)$ has arity $2$ in $\Oca_2$.
  Hence, $\varphi\left([\Product]_{I_1}\right)$ belongs to the line
  spanned by the binary generator of $\Oca_2$, that is there exists a
  scalar $\lambda\in\K$ such that $\varphi\left([\Product]_{I_1}\right)
  =\lambda[\Product]_{I_2}$. If there exists such a $\lambda$ different
  from zero, then for every nonzero scalar $\mu$, we have a well-defined
  operads morphism $\psi:\Oca_1\to\Oca_2$ satisfying
  $\psi([\Product]_{I_1}) =\mu[\Product]_{I_2}=
  \left(\mu/\lambda\right)\varphi([\Product]_{I_1})$. Hence,
  $\text{hom}\left(\Oca_1,\Oca_2\right)$ is either reduced to the zero
  morphism or it is in one-to-one correspondance with $\K$, which proves
  that $\text{hom}\left(\Oca_1,\Oca_2\right)$ is a vector space of
  dimension at most $1$. Moreover, if $\varphi$ is different from $0$,
  that is there is a nonzero scalar such that $\varphi([\Product]_{I_1})
  =\lambda[\Product]_{I_2}$, we have $\varphi(1/\lambda[\Product]_{I_1})
  =[\Product]_{I_2}$, so that $\varphi$ is surjective.
\end{proof}

We introduce the binary relation $\OrdKMag$ on $\QMag$ as follows: we
have $\Oca_2\OrdKMag\Oca_1$ if the dimension of
$\text{hom}\left(\Oca_1,\Oca_2\right)$ is equal to $1$.

\begin{Proposition} \label{prop:order_relations_on_QMag_and_ideals}
  Let $\Oca_1=\KMag/I_1$ and $\Oca_2=\KMag/I_2$ be two elements of
  $\QMag$.   We have $\Oca_2\OrdKMag\Oca_1$ if and only if
  $I_1\subseteq I_2$.
\end{Proposition}

\begin{proof}
  Since $\text{hom}\left(\Oca_1,\Oca_2\right)$ is a vector space of
  dimension at most $1$, it contains a nonzero morphism if and only if
  the morphism $\overline{\varphi}:\Oca_1\to\Oca_2$ satisfying
  $\overline{\varphi}([\Product]_{I_1})=[\Product]_{I_2}$ is well
  -defined, which means that $\Oca_2\OrdKMag\Oca_1$ is equivalent to this
  condition. Moreover, by the universal property of the quotient,
  $\overline{\varphi}$ is well-defined if and only if $I_1$ is included
  in the kernel of the morphism $\varphi:\KMag\to\Oca_2$ defined by
  $\varphi(\Product)=[\Product]_{I_1}$. This kernel is equal to
  $I_2$, so that $\overline{\varphi}$ is well-defined if and only if
  $I_1$ is included in $I_2$, which concludes the proof.
\end{proof}

@@

Given two elements $\Oca_1=\KMag/I_1$ and $\Oca_2=\KMag/I_2$ of $\QMag$,
we let
\begin{equation} \label{equ:definition_of_lattice_operations_of_QMag}
  \Oca_1\wedge \Oca_2=\KMag/\left(I_1+I_2\right)\ \ \text{and}\ \
  \Oca_1\wedge \Oca_2=\KMag/\left(I_1\cap I_2\right).
\end{equation}
Explicitely, for every integer $n$, $\left(\Oca_1\wedge\Oca_2\right)(n)$
(respectively, $\left(\Oca_1\vee\Oca_2\right)(n)$) is the quotient of
vector spaces $\KMag(n)/\left(I_1(n)+I_2(n)\right)$ (respectively,
$\KMag(n)/\left(I_1(n)\cap I_2(n)\right)$).

\begin{Theorem} \label{thm:lattice_structure_of_QMag}
  The tuple $\left(\QMag,\OrdKMag,\wedge,\vee\right)$ is a lattice.
\end{Theorem}

\begin{proof}
  The map $\mathcal{Q}:\IMag\to\QMag$ defined by $\mathcal{Q}(I)=\KMag/I$
  is a bijection by definition of $\QMag$. Moreover, from
  Proposition~\ref{prop:order_relations_on_QMag_and_ideals},
  $\Oca_2\OrdKMag\Oca_1$ is equivalent to $I_1\subseteq I_2$, so that
  $\left(\QMag,\OrdKMag\right)$ is a poset and $\mathcal{Q}$ is a
  decreasing bijection. The tuple $\left(\IMag,\subseteq,\cap,+\right)$
  being a lattice, the decreasing bijection $\mathcal{Q}$ induces
  lattice operations on $\QMag$, precisely $\wedge$ and $\vee$ by
  definition.
\end{proof}

@@

In Theorem~\ref{thm:Grassmann_formula_for__Hilbert_series_of_QMag}, we
show a formula, analogous to the Grassman formula, for Hilbert series of
elements of the lattice $\left(\QMag,\OrdKMag,\wedge,\vee\right)$. For
that, we need the following :

\begin{Lemma} \label{lem:lattice_structure_and_coefficients_of_Hiblert_series}
  Let $\Oca_1$ and $\Oca_2$ be two elements of $\QMag$. For every integer
  $n$, we have
  \begin{equation} \label{equ:lattice_structure_and_coefficients_of_the_Hiblert_series}
    \dim{\left(\Oca_1\wedge\Oca_2(n)\right)}+
    \dim{\left(\Oca_1\vee\Oca_2(n)\right)}=
    \dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}.
  \end{equation}
\end{Lemma}

\begin{proof}
  Let $I_1$ and $I_2$ be the two operadic ideals of $\KMag$ such that
  $\Oca_1$ and $\Oca_2$ are equal to $\KMag/I_1$ and $\KMag/I_2$,
  respectively. For every integer $n$, we have
  \[\begin{split}
  &\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}+
  \dim{\left(\Oca_1\vee\Oca_2(n)\right)}\\
  &=\dim{\Big(\KMag(n)/\left(I_1+I_2\right)(n)\Big)}+
  \dim{\Big(\KMag(n)/\left(I_1\cap I_2\right)(n)\Big)}\\
  &=\dim{\left(\KMag(n)\right)}-\dim{\left(I_1(n)+I_2(n)\right)}
  +\dim{\left(\KMag(n)\right)}-\dim{\left(I_1(n)\cap I_2(n)\right)}\\
  &=\dim{\left(\KMag(n)\right)}-\dim{\left(I_1(n)\right)}+
  \dim{\left(\KMag(n)\right)}-\dim{\left(I_2(n)\right)}\\
  &=\dim{\Big(\KMag(n)/I_1(n)\Big)}+\dim{\Big(\KMag(n)/I_2(n)\Big)}\\
  &=\dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}.
  \end{split}\]
  The third equality is due to the Grassmann formula applied to the
  subspaces $I_1(n)$ and $I_2(n)$ of $\KMag(n)$.
\end{proof}

\begin{Theorem} \label{thm:Grassmann_formula_for__Hilbert_series_of_QMag}
  Let $\Oca_1$ and $\Oca_2$ be two elements of $\QMag$. We have
  \begin{equation} \label{equ:lattice_structure_and_Hiblert_series}
    \HilbertSeries_{\Oca_1\wedge\Oca_2}(t)+\HilbertSeries_{\Oca_1\vee\Oca_2}(t)=\HilbertSeries_{\Oca_1}(t)+\HilbertSeries_{\Oca_2}(t).
    \end{equation}
  \end{Theorem}

\begin{proof}
  We have
  \[\begin{split}
  \HilbertSeries_{\Oca_1\wedge\Oca_2}(t)+\HilbertSeries_{\Oca_1\vee\Oca_2}(t)&=\sum_{n\in\N}\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_1\vee\Oca_2(n)\right)}t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1\wedge\Oca_2(n)\right)}+\dim{\left(\Oca_1\vee\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\dim{\left(\Oca_1(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_2(n)\right)}t^n\\
  &=\HilbertSeries_{\Oca_1}(t)+\HilbertSeries_{\Oca_2}(t).
  \end{split}\]
\end{proof}
The third equality is due to Lemma~\ref{lem:lattice_structure_and_coefficients_of_Hiblert_series}.
% Hilbert series, enumeration, a little of rewrite systems, etc.

>>>>

In this section, we equip the set of quotients of the linear magmatic
operad with a lattice structure. We also show the Grassmann formula
analog for this lattice.

@@

Throughout Section~\ref{sec:Magmatic_operads}, we fix a commutative field
$\K$. The \Def{linear magmatic operad}, written $\KMag$, is the free
linear operad over one binary generator. For each arity $n$, $\KMag(n)$
is the vector space with basis $\Mag(n)$, and the compositions of $\KMag$
extend by linearity the ones of $\Mag$.

We denote by $\IMag$ the set of operadic ideals of $\KMag$, and
\begin{equation} \label{equ:definition_of_QMag}
  \QMag =\Big\{\KMag/ I\mid I\in\IMag\Big\},
  \end{equation}
the set of quotients of the linear magmatic operad. Given two elements
$\Oca_1$ and $\Oca_2$ of $\QMag$, we denote by
$\text{hom}\left(\Oca_1,\Oca_2\right)$ the set of operads morphism from
$\Oca_1$ to $\Oca_2$.

@@

\begin{proof}
  The operad $\Oca_1$ is spanned by the binary generator
  $[\Product]_{I_1}$, which is the class of the element
  $\Product\in\KMag$ in $\Oca_1$. Hence, a morphism
  $\varphi:\Oca_1\to\Oca_2$ is uniquely determined by
  $\varphi\left([\Product]_{I_1}\right)$. Moreover,
  $\varphi\left([\Product]_{I_1}\right)$ has arity $2$ in $\Oca_2$.
  Hence, $\varphi\left([\Product]_{I_1}\right)$ belongs to the line
  spanned by the binary generator of $\Oca_2$, that is there exists a
  scalar $\lambda\in\K$ such that $\varphi\left([\Product]_{I_1}\right)
  =\lambda[\Product]_{I_2}$. If there exists such a $\lambda$ different
  from zero, then for every nonzero scalar $\mu$, we have a well-defined
  operad morphism $\psi:\Oca_1\to\Oca_2$ satisfying
  $\psi([\Product]_{I_1}) =\mu[\Product]_{I_2}=
  \left(\mu/\lambda\right)\varphi([\Product]_{I_1})$. Hence,
  $\text{hom}\left(\Oca_1,\Oca_2\right)$ is either reduced to the zero
  morphism or it is in one-to-one correspondance with $\K$, which proves
  that $\text{hom}\left(\Oca_1,\Oca_2\right)$ is a vector space of
  dimension at most $1$. Moreover, if $\varphi$ is different from $0$,
  that is there is a nonzero scalar such that $\varphi([\Product]_{I_1})
  =\lambda[\Product]_{I_2}$, we have $\varphi(1/\lambda[\Product]_{I_1})
  =[\Product]_{I_2}$, so that $\varphi$ is surjective.
\end{proof}

We introduce the binary relation $\OrdQMag$ on $\QMag$ as follows: we
have $\Oca_2\OrdQMag\Oca_1$ if the dimension of
$\text{hom}\left(\Oca_1,\Oca_2\right)$ is equal to $1$.

\begin{Proposition} \label{prop:order_relations_on_QMag_and_ideals}
  Let $\Oca_1=\KMag/I_1$ and $\Oca_2=\KMag/I_2$ be two elements of
  $\QMag$.   We have $\Oca_2\OrdQMag\Oca_1$ if and only if
  $I_1\subseteq I_2$.
\end{Proposition}

\begin{proof}
  Since $\text{hom}\left(\Oca_1,\Oca_2\right)$ is a vector space of
  dimension at most $1$, it contains a nonzero morphism if and only if
  the morphism $\overline{\varphi}:\Oca_1\to\Oca_2$ satisfying
  $\overline{\varphi}([\Product]_{I_1})=[\Product]_{I_2}$ is well
  -defined, which means that $\Oca_2\OrdQMag\Oca_1$ is equivalent to this
  condition. Moreover, by the universal property of the quotient,
  $\overline{\varphi}$ is well-defined if and only if $I_1$ is included
  in the kernel of the morphism $\varphi:\KMag\to\Oca_2$ defined by
  $\varphi(\Product)=[\Product]_{I_1}$. This kernel is equal to
  $I_2$, so that $\overline{\varphi}$ is well-defined if and only if
  $I_1$ is included in $I_2$, which concludes the proof.
\end{proof}

@@

Given two elements $\Oca_1=\KMag/I_1$ and $\Oca_2=\KMag/I_2$ of $\QMag$,
we let
\begin{equation} \label{equ:definition_of_lattice_operations_of_QMag}
  \Oca_1\InfQMag \Oca_2=\KMag/\left(I_1+I_2\right)\ \ \text{and}\ \
  \Oca_1\SupQMag \Oca_2=\KMag/\left(I_1\cap I_2\right).
\end{equation}
Explicitely, for every integer $n$, $\left(\Oca_1\InfQMag\Oca_2\right)(n)$
(respectively, $\left(\Oca_1\SupQMag\Oca_2\right)(n)$) is the quotient of
vector spaces $\KMag(n)/\left(I_1(n)+I_2(n)\right)$ (respectively,
$\KMag(n)/\left(I_1(n)\cap I_2(n)\right)$).

\begin{Theorem} \label{thm:lattice_structure_of_QMag}
  The tuple $\LatQMag$ is a lattice.
\end{Theorem}

\begin{proof}
  The map $\mathcal{Q}:\IMag\to\QMag$ defined by $\mathcal{Q}(I)=\KMag/I$
  is a bijection by definition of $\QMag$. Moreover, from
  Proposition~\ref{prop:order_relations_on_QMag_and_ideals},
  $\Oca_2\OrdQMag\Oca_1$ is equivalent to $I_1\subseteq I_2$, so that
  $\left(\QMag,\OrdQMag\right)$ is a poset and $\mathcal{Q}$ is a
  decreasing bijection. The tuple $\left(\IMag,\subseteq,\cap,+\right)$
  being a lattice, the decreasing bijection $\mathcal{Q}$ induces
  lattice operations on $\QMag$, precisely $\InfQMag$ and $\SupQMag$ by
  definition.
\end{proof}

@@

In order to show the Grassman formula analog for $\LatQMag$, we need the
following lemma.

\begin{Lemma} \label{lem:lattice_structure_and_coefficients_of_Hiblert_series}
  Let $\Oca_1$ and $\Oca_2$ be two elements of $\QMag$. For every integer
  $n$, we have
  \begin{equation} \label{equ:lattice_structure_and_coefficients_of_the_Hiblert_series}
    \dim{\left(\Oca_1\InfQMag\Oca_2(n)\right)}+
    \dim{\left(\Oca_1\SupQMag\Oca_2(n)\right)}=
    \dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}.
  \end{equation}
\end{Lemma}

\begin{proof}
  Let $I_1$ and $I_2$ be the two operadic ideals of $\KMag$ such that
  $\Oca_1$ and $\Oca_2$ are equal to $\KMag/I_1$ and $\KMag/I_2$,
  respectively. For every integer $n$, we have
  \[\begin{split}
  &\dim{\left(\Oca_1\InfQMag\Oca_2(n)\right)}+
  \dim{\left(\Oca_1\SupQMag\Oca_2(n)\right)}\\
  &=\dim{\Big(\KMag(n)/\left(I_1+I_2\right)(n)\Big)}+
  \dim{\Big(\KMag(n)/\left(I_1\cap I_2\right)(n)\Big)}\\
  &=\dim{\left(\KMag(n)\right)}-\dim{\left(I_1(n)+I_2(n)\right)}
  +\dim{\left(\KMag(n)\right)}-\dim{\left(I_1(n)\cap I_2(n)\right)}\\
  &=\dim{\left(\KMag(n)\right)}-\dim{\left(I_1(n)\right)}+
  \dim{\left(\KMag(n)\right)}-\dim{\left(I_2(n)\right)}\\
  &=\dim{\Big(\KMag(n)/I_1(n)\Big)}+\dim{\Big(\KMag(n)/I_2(n)\Big)}\\
  &=\dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}.
  \end{split}\]
  The third equality is due to the Grassmann formula applied to the
  subspaces $I_1(n)$ and $I_2(n)$ of $\KMag(n)$.
\end{proof}

The Grassman formula analog for $\LatQMag$ is given by the following
theorem.

\begin{Theorem} \label{thm:Grassmann_formula_for__Hilbert_series_of_QMag}
  Let $\Oca_1$ and $\Oca_2$ be two elements of $\QMag$. We have
  \begin{equation} \label{equ:lattice_structure_and_Hiblert_series}
    \HilbertSeries_{\Oca_1\InfQMag\Oca_2}(t)+\HilbertSeries_{\Oca_1\SupQMag\Oca_2}(t)=\HilbertSeries_{\Oca_1}(t)+\HilbertSeries_{\Oca_2}(t).
    \end{equation}
  \end{Theorem}

\begin{proof}
  We have
  \[\begin{split}
  \HilbertSeries_{\Oca_1\InfQMag\Oca_2}(t)+\HilbertSeries_{\Oca_1\SupQMag\Oca_2}(t)&=\sum_{n\in\N}\dim{\left(\Oca_1\InfQMag\Oca_2(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_1\SupQMag\Oca_2(n)\right)}t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1\InfQMag\Oca_2(n)\right)}+\dim{\left(\Oca_1\SupQMag\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\Big(\dim{\left(\Oca_1(n)\right)}+\dim{\left(\Oca_2(n)\right)}\Big)t^n\\
  &=\sum_{n\in\N}\dim{\left(\Oca_1(n)\right)}t^n+\sum_{n\in\N}\dim{\left(\Oca_2(n)\right)}t^n\\
  &=\HilbertSeries_{\Oca_1}(t)+\HilbertSeries_{\Oca_2}(t).
  \end{split}\]
\end{proof}
The third equality is due to Lemma~\ref{lem:lattice_structure_and_coefficients_of_Hiblert_series}.
% Hilbert series, enumeration, a little of rewrite systems, etc.
